---
title: "Strava Analysis"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

The idea of this project was primarily a way of putting into practice my learnings from Jad Khalili's Udemy Course on Git/GitHub (Git: Become an Expert in Git & GitHub in 4 Hours). I took this course in order to deepen my knowledge around best practices in data science. 

In this course Jad covers:

* An introduction to version/source control, Git and GitHub
* Theory around repositories, the 'staging area', and basic/advanced Git commands
* Git branches, and how to add, remove and merge them

The actual content of the project itself was a way to do some basic analysis of my Strava data. Strava is an online 'social-media' platform specifically desinged to connect athletes and allow them to share and record their exercise activities. 

To get access to my activity data I used the Strava API and the [rStrava package](https://github.com/fawda123/rStrava), and in order to create many of the visuals I used inspiration from [Marcus Volz](https://github.com/marcusvolz/strava).


## Grab your attention

```{r cars, echo=FALSE}
suppressPackageStartupMessages({
  library(tidyverse)
  library(yaml)
  library(httr)
  library(jsonlite)
  library(devtools)
  library(strava)
  library(rStrava)
  library(googlePolylines)
  library(gridExtra)
})

setwd('C:/Users/sma8lw/OneDrive - BP/Documents/Andrew/Other/strava-app')

data <- process_data('activities/')
p1 <- plot_facets(data)
p1
```

## Visualising my Runs

Strava members are also able to access their Strava data by requesting an archive of all of their activities [here](https://support.strava.com/hc/en-us/articles/216918437-Exporting-your-Data-and-Bulk-Export). Following these instructions allows you to get a mass export of all data, and most importantly it includes the actual GPS data files that contains latitude and longitude data. These files vary based on the method of recording your activities - my Samsung Galaxay Watch records them as GPX files.

This is how many of Marcus Volz's creative plots can be generated:

# Running frequency

You will notice there are two main periods that I have maintained consistent uploads to Strava:

1. Mid 2019 - in preparation for a marathon in Oct 2019 (dark red square)
2. Early 2020 - motivation to start running consistently again as Melbourne went into Covid lockdown

```{r runs, echo=FALSE}
p_cal <- plot_calendar(data, unit = "distance") + 
          ggtitle("Calendar of all runs") +
          theme(plot.title = element_text(hjust = 0.5))

p_cal

```

# Where have I been running?

The following plot shows my runs placed on a map, with the darkness of the lines indicating the number of times I have run over a particular part of Melbourne. Three things to note:

1. You'll notice there are two main parts of town that are darker than the rest. These represent when I moved house from Fitzroy North to Richmond.
2. The route at the bottom of the map is the marathon route that takes you down St Kilda.
3. The run I did where I spelt out my name (ANDY) :) 

```{r map, echo=FALSE}
p_map <- plot_map(data, lon_min = 144.863491, lon_max = 145.073007, lat_min = -37.883543, lat_max = -37.766141) + 
          ggtitle("Map of all my runs overlayed") +
          theme(plot.title = element_text(hjust = 0.5))
p_map
```
# What time of the day?

This visual by Marcus allowsyou to see the frequency of your runs by time of day and day of week.

```{r time of day, echo = F}
plot_ridges(data)

```

I thought it might be interesting to see how this distribution changed between last year and this COVID affected year where I have been working from home and have had the flexibility to run during the day. And sure enough, you can see that in the post-COVID plot, my runs have been much more dispersed across the day compared to pre-COVID.


```{r pre and post, echo = F}
pre_covid <- plot_ridges(data %>% filter(time <= "2020-03-01")) + ggtitle("Pre-Covid") +
          theme(plot.title = element_text(hjust = 0.5))
post_covid <- plot_ridges(data %>% filter(time > "2020-03-01")) + ggtitle("Post-Covid") +
          theme(plot.title = element_text(hjust = 0.5)) 

grid.arrange(pre_covid, post_covid, ncol = 2)

```

